---
title: "Lab Week 9 - Strings in stringr & other stuff"
author: "Sophia Leiker"
date: "3/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

We haven't done a ton of work with strings so far in 206 or 244. In this part, we'll learn a bit about how to deal with infamous **regular expressions** (*regex* or *regexp*) using nice functions in the `stringr` package (in the `tidyverse`) to make our lives easier.

Want to learn more about working with regex *without* the `stringr` package functions? Read a nice introduction from Roger Peng [HERE](https://bookdown.org/rdpeng/rprogdatascience/regular-expressions.html).

## Useful functions in `stringr`

We'll do some work with the `starwars` dataset from `dplyr` (so if you have the `tidyverse`) attached, the dataset already exists. 

Check it out by running `View(starwars)` in the Console. OK here we go. 

### Detect a pattern

Use `stringr::str_detect()` to detect a string pattern. 

Let's see how it works first, then try it out for something that might be useful: 

```{r}
# Make a little vector
vec <- c("  A Unicorn     Fish", " add   cornmeal ", "  French FRIES", " acorn squash  ")
 
# Then detect a string pattern: 
str_detect(vec, pattern = "corn") #give it a vector of character objects, such as a column in a dataframe, and it will tell you if it detects the specific strong that you are detecting for (results, in first one it found the pattern, second yes, third no, fourth yes)
str_detect(vec, pattern = "fish") #there is no elements with all lowercase!
str_extract(vec, pattern = "[a-z]") 
#the square brackets is asking for anything that falls between a-z
#returns n, a, r, a --> 
str_extract(vec, pattern = "[a-z]+")
#look for this pattern, the + tells it to look for it 1 or more times, * --> 0 or more times, it will stop when it hits a space or a capital letter
str_extract_all(vec, pattern = "[a-zA-Z]+")
#this is lower case a to lowercase z and upper case a to upper case Z
```


We see that it returns a T/F vector indicating whether or not the pattern "corn" exists within each element of the vector. 

How might we use that in a data frame? It can be helpful to use `str_detect()` within `filter()` to only keep observations that contain that pattern in the specified variable. 

For example: Only keep observations where the pattern "walk" exists within the character name. 

```{r}
walk_subset <- starwars %>% 
  filter(str_detect(name, pattern = "walk")) #detecting only names which includes walk in it
```


### Replace a pattern

Use `stringr::str_replace()` to replace a pattern with something else.

Let's start with our vector, `vec`, replacing "corn" with "COOL":
```{r}
str_replace(vec, pattern = "corn", replacement = "COOL") 
#if you detect "corn" replace it with something, in this case "COOL", this will only replace the first instance it sees it for each of the items in the vector
 
str_replace_all(vec, pattern = "a", replacement = "ZZZ")
#this will replace it every time it shows up
```

And with the `starwars` data, let's replace any "Sky" string pattern with "TORNADO": 

```{r}
sw_tornado <- starwars %>% 
  mutate(name = str_replace(name, pattern = "Sky", replacement = "TORNADO"))
```


### Remove white space

In some of the strings of our vector `vec`, we have really obnoxious white space. This can be problematic because excess white spaces can mess with groupings or finding matches (i.e. because " hello friend" is, in R's mind, definitely not the same group as "hello friend" or "hello   friend ").

There are a few helper functions in the `stringr` package to help us out. Let's use two of them: `str_trim()` and `str_squish()`. 

- `str_trim()`: removes leading and trailing whitespace
- `str_squish()`: removes leading / trailing whitespace and excess interior whitespace

```{r}
str_trim(vec) # Notice there are still excess spaces in the middle
#this takes out the leading white space
 
str_squish(vec) # Notice that now trailing, leading and excess interior white spaces are removed
```

### Change string case

It can also be useful to update the *case* of strings in your code. Since, for example, "Fish" will be seen entirely differently from "fish". 
We can use the `str_to_*()` functions to change the case. 

- `str_to_lower()`: make everything lowercase
- `str_to_upper()`: make everything to uppercase
- `str_to_title()`: convert to title case

Let's try it out with our vector: 

```{r}
str_to_lower(vec) #all lowercase
 
str_to_upper(vec) #all uppercase
 
str_to_title(vec) #title case
```

And we can use these things in combination, usually within `mutate()` if it's to update a column in a data frame.

For example, with `starwars` data: 

```{r}
sw_combo <- starwars %>% 
  mutate(name = str_to_lower(name)) %>% #all lowercase
  filter(str_detect(name, pattern = "darth")) %>% #in the name column, detect darth
  mutate(name = str_replace(name, pattern = "darth", replacement = "Llama")) #replace darth with llama

```

### End `stringr` intro



