---
title: "Lab Week 9 - Strings in stringr & other stuff"
author: "Sophia Leiker"
date: "3/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

We haven't done a ton of work with strings so far in 206 or 244. In this part, we'll learn a bit about how to deal with infamous **regular expressions** (*regex* or *regexp*) using nice functions in the `stringr` package (in the `tidyverse`) to make our lives easier.

Want to learn more about working with regex *without* the `stringr` package functions? Read a nice introduction from Roger Peng [HERE](https://bookdown.org/rdpeng/rprogdatascience/regular-expressions.html).

## Useful functions in `stringr`

We'll do some work with the `starwars` dataset from `dplyr` (so if you have the `tidyverse`) attached, the dataset already exists. 

Check it out by running `View(starwars)` in the Console. OK here we go. 

### Detect a pattern

Use `stringr::str_detect()` to detect a string pattern. 

Let's see how it works first, then try it out for something that might be useful: 

```{r}
# Make a little vector
vec <- c("  A Unicorn     Fish", " add   cornmeal ", "  French FRIES", " acorn squash  ")
 
# Then detect a string pattern: 
str_detect(vec, pattern = "corn")
str_detect(vec, pattern = "fish")
str_extract(vec, pattern = "[a-z]")
str_extract(vec, pattern = "[a-z]+")
str_extract_all(vec, pattern = "[a-zA-Z]+")
```

